---
title: "Income Change by industry"
format: html
editor: visual
---

```{r}
library(rio)
library(tidyverse)
library(vtable)
library(dplyr)
library(lubridate)
library(fixest)
library(ggplot2)
library(nnet)
```

```{r}
data_finance_finc <- read.csv('Retail_FIRR_Faminc.csv')

data_finance_finc <- data_finance_finc%>%
  filter(EMPSTAT == '10')

data_info_finc <- read.csv('Retail_Info_Faminc.csv')

data_info_finc <- data_info_finc %>%
  filter(EMPSTAT =='10')


```

```{r}
cutoff_date <- ('2020-03-01')

data_finance_finc <- data_finance_finc%>%
  mutate(covid = data_finance_finc$date > cutoff_date)

data_info_finc <- data_info_finc%>%
  mutate(covid = data_info_finc$date > cutoff_date)

data <- data_info_finc%>%
  filter(!(IND %in% (c('6470', '6480', '6570', '6590', '6670', '6780'))))%>%
  filter(!(FAMINC%in% (c('995','996','997','998','999'))))

data_finance <- data_finance_finc%>%
  filter(!(IND %in% (c('6470', '6480', '6570', '6590', '6670', '6780', '7080', '7181', '7190'))))%>%
  filter(!(FAMINC%in% (c('995','996','997','998','999'))))
```

```{r}

#mutate func
data_finance_finc <- data_finance_finc %>%
  group_by(FAMINC) %>%
  mutate(Low = if_else(FAMINC < 720, 1, 0)) %>%
  mutate(Middle = if_else(FAMINC > 720 & FAMINC < 840, 1, 0)) %>%
  mutate(High = if_else(FAMINC > 840 & FAMINC < 843, 1, 0))

data_info_finc <- data_info_finc %>%
  group_by(FAMINC) %>%
  mutate(Low = if_else(FAMINC < 720, 1, 0)) %>%
  mutate(Middle = if_else(FAMINC > 720 & FAMINC < 840, 1, 0)) %>%
  mutate(High = if_else(FAMINC > 840 & FAMINC < 843, 1, 0))



```

```{r}
reg_low_info <- 
  feols(Low ~ covid + indname + covid*indname , data = data_info_finc)
reg_mid_info<-
  feols(Middle ~ covid + indname + covid*indname , data = data_info_finc)

reg_high_info<-
  feols(High ~ covid + indname + covid*indname , data = data_info_finc)

etable(reg_low_info, reg_mid_info, reg_high_info)

```

In the presence of covid the retail employees are .4% lower chance of being in the low income bracket compared to the information industry. Low income here is defined as a family income less than \$30,000.

In the presence of covid the retail employees are .02% higher chance of being in the middle income bracket compared to the information industry. Middle income here is defined as a family income between \$30,000 to \$74,999.

In the presence of covid the retail employees are 3.9% higher chance of being in the high income bracket compared to the information industry. High income here is defined as a family income more than \$75,000.

To conclude with the presence of covid on employees wages was modest but positive toward the retail industry when compared to the information industry. We can conclude that employees wages grew more in retail than information.

```{r}
reg_low_finance <- feols(Low ~ covid + indname + covid*indname, data = data_finance_finc)
reg_mid_finance <- feols(Middle ~ covid + indname + covid*indname, data = data_finance_finc)
reg_high_finance <- feols(High ~ covid + indname + covid*indname, data = data_finance_finc)
etable(reg_low_finance, reg_mid_finance, reg_high_finance)
```

In the presence of covid the retail employees are 1.3% lower chance of being in the low income bracket compared to the Finance Industry. Low income here is defined as a family income less than \$30,000.

In the presence of covid the retail employees are 1.0% higher chance of being in the middle income bracket compared to the Finance Industry. Middle income here is defined as a family income between \$30,000 to \$74,999.

In the presence of covid the retail employees are 1.6% higher chance of being in the middle income bracket compared to the information industry. Middle income here is defined as a family income greater than \$75,000.
